---
title: "New Credible Set: `corrected_cs`"
author: "Anna Hutchinson"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, set.seed = 3, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(3)
library(corrcoverage)
```

```{r echo = FALSE}
prop_cov <- function(x) {
    sum(x$covered)/length(x$covered)
}
```

This vignette will show users how the `corrcoverage` package can be used to obtain a new credible set of variants that contains the true causal variant with some specified desired coverage value whilst containing as few variants as possible.

---

## Simulate GWAS data

As in the [corrected coverage vignette](https://annahutch.github.io/corrcoverage/articles/my-vignette.html), let's begin by simulating some GWAS data using the `simGWAS` package.

```{r}
library(simGWAS)

#  Simulate reference haplotypes
nsnps <- 200
nhaps <- 1000
lag <- 30  # genotypes are correlated between neighbouring variants
maf <- runif(nsnps + lag, 0.05, 0.5)  # common SNPs
laghaps <- do.call("cbind", lapply(maf, function(f) rbinom(nhaps, 1, f)))
haps <- laghaps[, 1:nsnps]
for (j in 1:lag) haps <- haps + laghaps[, (1:nsnps) + j]
haps <- round(haps/matrix(apply(haps, 2, max), nhaps, nsnps, byrow = TRUE))
snps <- colnames(haps) <- paste0("s", 1:nsnps)
freq <- as.data.frame(haps + 1)
freq$Probability <- 1/nrow(freq)
sum(freq$Probability)
MAF <- colMeans(freq[, snps] - 1)  # minor allele frequencies
CV <- sample(snps[which(colMeans(haps) > 0.1)], 1)
iCV <- sub("s", "", CV)  # index of cv
LD <- cor2(haps) # correlation between SNPs
```

```{r}
OR <- 1.1 # odds ratios
N0 <- 10000 # number of controls
N1 <- 10000 # number of cases
  
z0 <- simulated_z_score(N0 = N0, # number of controls
                        N1 = N1, # number of cases
                        snps = snps, # column names in freq of SNPs for which Z scores should be generated
                        W = CV, # causal variants, subset of snps
                        gamma.W = log(OR), # log odds ratios
                        freq = freq) # reference haplotypes
```

To calculate $V$, the prior variance for the estimated effect size, we use `coloc::Var.data.cc`.

```{r}
varbeta <- coloc:::Var.data.cc(f = MAF, N = N1+N0, s = N1/(N0+N1)) # variance of estimated effect size
```

We can then use the `ppfunc` function to calculate the posterior probabilities of causality for each variant.

```{r}
postprobs <- ppfunc(z = z0, V = varbeta) 
```

We use the `est_mu` function to obtain an estimate of the true effect at the causal variant.

```{r}
muhat <- est_mu(z0, MAF, N0, N1)
muhat
```

---

## Corrected Coverage Estimate

The `corrected_cov` function is used to find the corrected coverage of a credible set with specified threshold, say 0.9.

Note that this function is similar to using `corrcov` as explained in the previous vignette; which would require $Z$-scores, minor allele frequencies and sample sizes. Here, we have already calculated some of the intermediaries calculated in the `corrcov` function (muhat, varbeta etc.) so we can use `corrected_cov` instead.

```{r}
thr = 0.9
corrcov <- corrected_cov(mu = muhat, V = varbeta, Sigma = LD, pp0 = postprobs, thr = thr, nrep = 1000)
cs <- credset(pp = postprobs, thr = thr)
data.frame(claimed.cov = cs$claimed.cov, corr.cov =  corrcov, nvar = cs$nvar)
```

Using the Bayesian approach for statistical fine-mapping we obtain a 90% credible set consisting of 2 variants. The claimed coverage of this credible set is ~0.90, yet the corrected coverage estimate is ~0.96, suggesting that we can afford to be 'more confident' that we have captured the causal variant in our credible set. 

## Empirical Coverage Estimate

Again, for the purpose of this vignette we can investigate how accurate this estimate is by simulating many credible sets from the same system, and finding the proportion of these that contain the true causal variant.

```{r}
z0.tmp <- simulated_z_score(N0 = N0, # number of controls
                            N1 = N1, # number of cases
                            snps = snps, # column names in freq of SNPs for which Z scores should be generated
                            W = CV, # causal variants, subset of snps
                            gamma.W = log(OR), # log odds ratios
                            freq = freq, # reference haplotypes
                            nrep = 5000) 

pps <- ppfunc.mat(zstar = z0.tmp, V = varbeta)  # find pps from a matrix of simulations
cs.cov <- apply(pps, 1, function(x) credset(x, CV = iCV, thr = thr)$cov)
true.cov.est <- mean(cs.cov)
data.frame(claimed.cov = cs$claimed.cov, corr.cov =  corrcov, true.cov = true.cov.est, nvar = cs$nvar)
```

We see that our corrected coverage value of ~0.962 is a good estimate of the empirical coverage estimate, ~0.958.

---

## Obtain a New Credible Set

Our results suggest that we may be able to remove some variants from the credible set, whilst still achieving the desired coverage of 90%.

The `corrected_cs` function uses GWAS summary statistics and some user-defined parameters to find the smallest credible set such that the coverage estimate is within some accuracy of the desired coverage.

The function requires the $Z$-scores (`z`), minor allele frequencies (`f`), control and case sample sizes (`N0` and `N1` respectively), LD matrix (`Sigma`), some lower threshold value (`lower`), some upper threshold value (`upper`) and the desired coverage (`desired.cov`). 

The function uses the [bisection root finding method](https://en.wikipedia.org/wiki/Root-finding_algorithm#Bisection_method) to converge to the smallest threshold such that the corrected coverage is larger than the desired coverage. The accuracy parameter is optional (default value set to 0.005) and controls how much greater than the desired coverage the user is happy for the corrected coverage value to lie. 

The function reports the threshold values tested and their corresponding corrected coverage value. The maximum number of iterations for the bisecting root finding algorithm is an optional parameter, with default value 20. The functions stops when either the number of iterations reaches the maximum, or the corrected coverage is within some accuracy of the desired coverage. 

```{r eval = FALSE}
res <- corrected_cs(z = z0, f = MAF, N0, N1, Sigma = LD, lower = 0.5, upper = 1, desired.cov = 0.9)
res
```

In this example we see that a threshold of just 0.75 is required to obtain a credible set with 90% coverage of the causal variant. In the standard Bayesian approach, a threshold of 90% is chosen which we have seen leads to over-coverage.

The output is a credible set containing only 1 variant, that has a corrected coverage estimate of 0.9. This suggests that if the user implemented the standard Bayesian approach to statistical fine-mapping, then their 90% credible set would contain 2 variants and actually have nearer to 96% probability of containing the true causal variant. 

By applying the `corrected_cs` function, the user can obtain a new 90% credible set that has 90% probability of containing the true causal variant. This credible set contains only 1 variant, showing that the resolution of the fine-mapping experiment has improved. 

Finally, we can compare this coverage estimate of the corrected credible set to an empirical estimate of the true coverage. 

```{r eval = FALSE}
new.cs.sims <- apply(pps, 1, function(x) credset(x, CV = iCV, thr = res$req.thr)$cov)
true.cov.est2 <- mean(new.cs.sims)
```

Original 90% credible set:
```{r echo=FALSE, eval = FALSE}
df1 <- data.frame(claimed.cov = round(cs$claimed.cov, 3), corr.cov =  round(corrcov, 3), true.cov = round(true.cov.est, 3), nvar = cs$nvar)
print(df1, row.names = FALSE)
```

New 90% credible set:
```{r echo=FALSE, eval = FALSE}
df2 <- data.frame(claimed.cov = round(res$size, 3), corr.cov = round(res$corr.cov, 3), true.cov = round(true.cov.est2, 3), nvar = length(res$credset))
print(df2, row.names = FALSE)
```

---

This vignette has shown 2 things:

1. The 90% credible set obtained using the standard Bayesian approach for statistical fine-mapping has nearer to 96% coverage of the causal variant.

2. If we did indeed want a credible set with 90% coverage of the causal variant, then we can use a threshold of 0.75 in the Bayesian approach to fine-mapping, resulting in a credible set containing only 1 variant. 

